The growth of children's semantic and phonological networks: insight from 10 languages

### 2019-02-12 journal/ms.tex
final version

  \abstract{Children tend to produce words earlier when they are connected to a
variety of other words along the phonological and semantic dimensions.
Though this connectivity effect has been extensively documented, little
is known about the underlying developmental mechanism. One view suggests
that learning is primarily predicted by a network growth model where
highly connected words in the child's early lexicon enable learning of
similar words. Another view suggests that learning is driven by highly
connected words in the external learning environment, instead of highly
connected words in the early internal lexicon. The present study tests
both scenarios systematically in both the phonological and semantic
domains across 10 languages. We show that external connectivity in the
learning environment drives growth in both production- and
comprehension-based vocabularies, even controlling for word frequency
and length. This pattern of findings suggests a word learning mechanism
where children harness their statistical learning abilities to detect
and learn highly connected words in the learning environment.}
  \keywords{Word learning; semantic network; phonological network; network growth;
cross-linguistic analysis. \\

@@

What factors shape vocabulary learning over the course of early
childhood? To investigate this question, scientists have adopted
multiple research strategies, from conducting controlled laboratory
experiments (e.g. Markman, 1990) to analyzing dense corpora capturing
language learning in context (e.g., B. C. Roy, Frank, DeCamp, Miller, \&
Roy, 2015). One prominent strategy consists in documenting the timeline
of words' acquisition, and studying the properties that make words easy
or hard to learn (e.g., J. C. Goodman, Dale, \& Li, 2008; Huttenlocher,
Haight, Bryk, Seltzer, \& Lyons, 1991). For example, J. C. Goodman et
al. (2008) found that, within a lexical category (e.g., nouns), higher
parental frequency is associated with earlier learning. Researchers have
studied the role of several other factors such as word length and the
mean length of utterances in which the word occurs (e.g., Braginsky,
Yurovsky, Marchman, \& Frank, 2016; Swingley \& Humphrey, 2018).

Besides word-level properties, the structure of the lexicon (that is,
how words relate to one another) was also found to predict the Age of
Acquisition (AoA) of words. The lexical structure can be characterized
in terms of a network where each node represents a word in the
vocabulary, and each link between two nodes represents a relationship
between the corresponding pair of words (e.g., Collins \& Loftus, 1975;
Luce \& Pisoni, 1998). Previous studies have investigated early
vocabulary structure by constructing networks using a variety of
word-word relations including shared semantic features (McRae, Cree,
Seidenberg, \& McNorgan, 2005), target-cue relationships in free
association norms (Nelson, McEvoy, \& Schreiber, 1998), co-occurrence in
child directed speech (MacWhinney, 2014), and phonological relatedness
(Vitevitch, 2008). These studies have generally found that children tend
to produce nouns that have higher neighborhood density (i.e., high
connectivity in the network) earlier, both at the phonological and the
semantic level (Carlson, Sonderegger, \& Bane, 2014; Hills, Maouene,
Riordan, \& Smith, 2010; Hills, Maouene, Maouene, Sheya, \& Smith, 2009;
Stella, Beckage, \& Brede, 2017; Storkel, 2009).

@@

\caption{Illustration of the two growth scenarios. Filled grey circles represent known words (Internal) at a certain point in time. The empty, numbered circles represent words that have not yet been learned (External) and which are candidates to enter the lexicon next. The identity of the word that is going to be learned depends on the growth scenario.  Here the squares indicate the node that drives growth in each scenario and the hand pointer indicates which word is likely to be learned. For INT, the utility of a candidate, external node is the average degree (i.e., number of links) of the internal nodes that it would attach to. Thus, according to INT, node 1 is more likely to enter the lexicon. For EXT, the utility of a candidate node is its degree in the entire network. According to EXT, node 2 is more likely to enter the lexicon next.}\label{fig:growth}
\end{figure}

Hills et al. (2009) investigated the growth of lexico-semantic networks
in toddlers and found that growth did not proceed according to INT as
was originally hypothesized by Steyvers and Tenenbaum (2005), but rather
according to
EXT.\footnote{Besides INT and EXT, the authors tested a third mechanism (called the lure of associates) which resembles EXT in that it is driven by the connectivity of external nodes, except that this connectivity is computed with respect to words that are known. However, EXT is the externally-driven scenario that best predicted the data in this previous work.}
This is an important finding because it suggests that learning in the
early stages is mostly driven by properties of the external input,
regardless of how past knowledge is organized. However, this work has
been done in the special case of networks that are based on 1) semantic
associations, 2) production-based vocabularies, and 3) data from
English-learning children, only. The extent to which this result depends
on the domain (e.g., semantic vs.~phonological connectivity), the
vocabulary measure (production vs.~comprehension) and culture/language
is still unclear. In this work, we test the generality of the finding
along these three dimensions.

First, we study the phonological network in addition to the semantic
network. These two networks represent different ways the mental lexicon
is structured (Beckage \& Colunga, 2016). In particular, words that are
neighbors in the semantic network (e.g., \enquote{cat}, \enquote{dog})
are not necessarily neighbors in the phonological network, and vice
versa. Does the phonological network also predict word learning?
Previous work did find an effect of nouns' connectivity in the
phonological network on their age of learning (Carlson et al., 2014;
Stella et al., 2017; Storkel, 2009). In other words, nouns learned
earlier in life tend to sound similar to many other nouns than a word
learned later in life. However, this finding is \emph{a priori}
compatible with both INT and EXT, and previous studies did not
explicitly compare these two mechanisms. Here, we investigate whether
phonological networks, like semantic networks, grow through EXT, or if
they rather grow via INT (Figure~\ref{fig:growth}).

Second, we study vocabularies measured using both comprehension and
production. Previous studies have found differences between these
vocabularies in terms of their content and rate of acquisition (Bates,
Dale, \& Thal, 1995; Benedict, 1979; Fenson et al., 1994). These
differences may reflect the fact that comprehension and production do
not share the same constraints. For instance, whereas comprehension
depends on the ease with which words are stored and accessed, production
depends, additionally, on the ease with which words are articulated,
e.g., shorter words are produced earlier (Braginsky et al., 2016). By
investigating comprehension-based vocabularies, we assess the extent to
which the network growth mechanism captures general learning patterns
beyond the specific constraints of production.

Finally, we use developmental data in 10 languages. Lexical networks can
show more or less cross-linguistic variability along both the semantic
and phonological domains (Arbesman, Strogatz, \& Vitevitch, 2010; Lupyan
\& Lewis, 2017; Youn et al., 2016). Besides, cultures might differ in
the way caregivers talk to children (Cristia, Dupoux, Gurven, \&
Stieglitz, 2017; Kuhl et al., 1997), and this difference in the input
could influence the way in which the children's networks grow. Thus,
cross-linguistic comparison is crucial to test what growth mechanism is
cognitively universal and is used by all children, and what mechanism is
specific to some patterns of learning that emerge due to the particulars
of a given language or culture (Bates \& MacWhinney, 1987; Slobin,
2014).

We adopted the following research strategy. We used a cross-linguistic
dataset of children's vocabulary development (Fenson et al., 1994;
Frank, Braginsky, Yurovsky, \& Marchman, 2017), and we studied the
timeline of noun learning using the normative age of acquisition (i.e.,
the age at which at least 50\% of children know a given words). Our
choice of studying the normative learning trajectory rather than the
individual trajectories was motivated by the nature of the dataset
used---which is primarily based on cross-sectional studies. Children may
vary in their individual learning trajectories, but the aggregate data
provide highly robust measures of the average learning patterns (Fenson
et al., 1994). The use of such measures has lead to important insights
on the mechanisms of word learning (Braginsky et al., 2016; J. C.
Goodman et al., 2008; Hills et al., 2010, 2009; Stella et al., 2017;
Storkel, 2009).

The paper is organized as follows. First, we describe the datasets we
used and explain how we constructed the networks. Second, we analyze
static properties of words' connectivity in these networks (correlation
with age of acquisition and shape of the distribution) and we explain
how these properties inform hypotheses about network growth. Next, we
explicitly fit the two hypothesized growth mechanisms to the data. We
investigate the extent to which the results obtained in Hills et al.
(2009) generalize to phonological networks and comprehension-based
vocabularies, and whether this generalization holds
cross-linguistically.

@@

For each word in the CDI data, we compute the proportion of children who
understand or produce the word at each month. Then we fit a logistic
curve to these proportions and determined when the curve crosses 0.5,
i.e., the age at which at least 50\% of children know the word. We take
this point in time to be each word's age of acquisition (Braginsky et
al., 2016; J. C. Goodman et al., 2008).

@@

\caption{\label{tab:stats}Statistics for the dataset we used.}
\centering
\begin{tabular}[t]{>{\bfseries}lrlrl}
\toprule
\multicolumn{1}{c}{} & \multicolumn{2}{c}{Comprehension} & \multicolumn{2}{c}{Production} \\
\cmidrule(l{2pt}r{2pt}){2-3} \cmidrule(l{2pt}r{2pt}){4-5}
Language & Nouns & Ages & Nouns & Ages\\
\midrule
Croatian & 209 & 8-16 & 312 & 16-30\\
Danish & 200 & 8-20 & 316 & 16-36\\
English & 209 & 8-18 & 312 & 16-30\\
French & 197 & 8-16 & 307 & 16-30\\
Italian & 209 & 7-24 & 312 & 18-36\\
Norwegian & 193 & 8-20 & 316 & 16-36\\
Russian & 207 & 8-18 & 314 & 18-36\\
Spanish & 208 & 8-18 & 312 & 16-30\\
Swedish & 205 & 8-16 & 339 & 16-28\\
Turkish & 180 & 8-16 & 297 & 16-36\\
\bottomrule
\end{tabular}
\end{table}

@@

Since the free association norms are available only in English, we used
the hand-checked translation equivalents available in Wordbank, which
allowed us to use the English association norms across 10 languages.
Using the same association data across languages does not necessarily
lead to similar networks. Indeed, though this approximation assumes that
the semantic similarity measure is universal---which is a reasonable
approximation (e.g., Youn et al., 2016)---the set of words acquired by
children as well as the timeline of this acquisition can still vary from
language to language leading to possibility different learning
strategies.

@@

In previous studies, two nodes were linked if they had an edit distance
of 1 (Carlson et al., 2014; Stella et al., 2017; Storkel, 2009).
However, these studies reported a contribution of phonological
connectivity to noun learning when networks were built using a rich
adult vocabulary. Since the focus of the current study is on the
mechanism of growth, the networks should be based on children's early
vocabulary. The latter, however, contains very few noun pairs with an
edit distance of 1. Thus, we increased the threshold from 1 to 2, that
is, two nodes were related if their edit distance was equal to 1 or
2.\footnote{We also considered the case of an edit distance of 1 as well as the continuous measure, i.e., the inverse edit distance without threshold. In both cases, the results were weaker than those obtained with a threshold of 2. We did not consider the case of a threshold larger than 2 since many pairs become phonologically unrelated.}
The connectivity of a given node was characterized with its
\emph{degree}: the number of links it shares with other words.

@@

We start by analyzing word connectivity in the global (static) network.
We constructed this network using nouns learned by the oldest age for
which we have CDI data (e.g., in English this corresponds, in
comprehension, to the network by 18 months, and in production, to the
network by 30 months). This global network is the end-state towards
which both INT and EXT converge by the last month of learning. Moreover,
following Hills et al. (2009), we used this end-state network as a proxy
for the external connectivity in the learning environment. Below we
analyze properties of this global networks that may a priori hint at an
INT- or EXT-like growth.

\begin{figure}[!h]
\includegraphics[width=\textwidth]{ms_files/figure-latex/corrProd-1} \caption{Age of production in the global network as predicted by the degree (i.e., connectivity) in this network. Results are shown in each language for phonological and semantic networks. Each point is a word, with lines indicating linear model fits, and numbers indicating the Pearson correlation coefficients.}\label{fig:corrProd}
\end{figure}

\begin{figure}[!h]
\includegraphics[width=\textwidth]{ms_files/figure-latex/corrComp-1} \caption{Age of comprehension in the global network as predicted by the degree (i.e., connectivity) in this network. Results are shown in each language for phonological and semantic networks. Each point is a word, with lines indicating linear model fits, and numbers indicating the Pearson correlation coefficients.}\label{fig:corrComp}
\end{figure}

@@

Figures \ref{fig:corrProd} and \ref{fig:corrComp} show how the age of
production and comprehension, respectively, correlates with the degree
(or indegree for the semantic networks). For ease of visual comparison,
the predictor (i.e., the degree) was centered and scaled. The plots
show, overall, a negative correlation between the month of acquisition
and the degree. In production data, the average correlation across
languages was -0.24 (\(SD=\) 0.10) for the semantic networks and -0.30
(\(SD=\) 0.05) for the phonological networks. In comprehension data, the
average correlation was -0.21 (\(SD=\) 0.08) for the semantic networks
and -0.21 (\(SD=\) 0.07) for the phonological networks. These results
indicate that nouns with higher degrees are generally learned earlier,
thus replicating previous findings in English (Hills et al., 2009;
Storkel, 2009) and extending these findings to 10 different languages,
generally, in both production- and comprehension-based vocabularies.

@@

We estimated the parameter \(\beta\) using a Bayesian approach. The
inference was performed using the probabilistic programming language
WebPPL (N. Goodman \& Stuhlmuller, 2014). We defined a uniform prior
over \(\beta\), and at each month, we computed the likelihood function
over words that could possibly enter the lexicon at that month, fit to
the words that have been learned at that month (using formula 1). Markov
Chain Monte Carlo sampling resulted in a posterior distribution over
\(\beta\), which we summarized in Figure \ref{fig:growthPred}. The
results replicate Hills et al.'s original finding regarding the semantic
network in English and the production-based vocabulary, which is that
this network grows by EXT, not by INT. Crucially, our results show that,
generally speaking, this finding generalizes to comprehension-based
vocabulary, and holds across languages. This generalization was obtained
in both the semantic\footnote{One could imagine that the fact of using
  English free association norms cross-linguistically would decrease the
  effect of non-English semantic networks because of possible cultural
  differences. However, our findings do not support this assumption,
  rather, it supports our initial approximation about the universality
  of the semantic similarity measure. That said, this approximation is
  not perfect. For example there is evidence that a small part of the
  variance in free association data can be explained by phonological
  similarity (Kachergis, Cox, \& Jones, 2011; Matusevych \& Stevenson,
  2018), thus leading to possibly minor cross-linguistic differences.}
and phonological domains.

@@

Above we showed that the way semantic and phonological information is
structured in the learning environment contributes to noun learning (via
EXT) across languages. However, we know that other factors influence
learning as well (e.g., Braginsky et al., 2016). Next we investigated
how semantic and phonological connectivity interact with two other
factors. The first one is word frequency, a well studied factor shown to
predict the age of acquisition in a reliable fashion (e.g. J. C. Goodman
et al., 2008). The second factor is word length, which was shown to
correlate with phonological connectivity: Shorter words are more likely
to have higher connectivity (D. B. Pisoni, Nusbaum, Luce, \& Slowiaczek,
1985; Vitevitch \& Rodríguez, 2005).

@@

For word length, we counted the number of phonemes in our generated IPA
transcription. For word frequency, we used the frequency estimates from
Braginsky et al. (2016) where unigram counts were derived based on
CHILDES corpora in each language (MacWhinney, 2014). Although these
frequency counts use transcripts from independent sets of children, they
are based on large samples, and this allows us to average out possible
differences between children and the specificities of their input (see
J. C. Goodman et al., 2008 for a similar research strategy).

@@

The findings for the new predictors were as follows. Overall, frequency
is the largest and most consistent predictor of age of acquisition in
both comprehension and production data and across languages, endorsing
results for nouns across a variety of analyses (Braginsky et al., 2016;
J. C. Goodman et al., 2008; B. C. Roy et al., 2015). Word length is more
predictive for production than comprehension (and this difference is
very clear in the global model), replicating previous work (Braginsky et
al., under review). Thus, word length seems to reflect the effects of
production's constraints rather than comprehension's constraints, i.e.,
longer words are harder to articulate but they may not be significantly
more difficult to store and access.

@@

We found that the original findings reported in Hills et al. (2009)
generalize well across all these dimensions. First, just like semantic
networks, phonological networks grow via the externally-driven scenario
(EXT), not by the internally-driven mechanism (INT). Second,
comprehension-based vocabularies grow in a way similar to
production-based vocabularies. Finally, the findings were, overall,
similar across the 10 languages we tested. Although we find some
cross-linguistic variation when semantic and phonological networks were
pitted against frequency and length, this variability is to be taken
with a grain of salt as it might be exaggerated in our study by the
limited and partially-overlapping sample of nouns for each language. In
fact, both phonological and semantic connectivity are significant
predictors above and beyond frequency and length when data are pooled
across languages.

@@

For example, in the semantic domain, growth according to EXT can be
explained by a mechanism similar to cross-situational learning (McMurray
et al., 2012; Smith \& Yu, 2008; Yurovsky \& Frank, 2015). According to
this mechanism, children track the co-occurrence of concrete nouns with
their possible semantic referents. The referent of a word heard in only
one naming situation can be ambiguous (e.g., when the word
\enquote{ball} is heard for the first time in the presence of both a
ball and a chair), but hearing the same word in a diversity of semantic
contexts allows the learner to narrow down the set of possible
word-object mappings. In our case, free association (used to determine
semantic network connectivity) is related to contextual co-occurrence
(Fourtassi \& Dupoux, 2013; Griffiths, Steyvers, \& Tenenbaum, 2007),
meaning that highly connected words will tend to occur in a variety of
speech and referential contexts. This fact makes such words easier to
learn because they have more referential disambiguating cues across
learning contexts. Crucially, children can learn these words without
necessarily knowing the meaning of all other words with witch they
co-occur (hence the similarity with EXT). This possibility is supported
by the finding that words' diversity of occurrence in child directed
speech predicts their age of learning (Hills et al., 2010; Stella et
al., 2017).

@@

@@

>>>>

  \abstract{Children tend to produce words earlier when they are connected to a
variety of other words along the phonological and semantic dimensions.
Though these semantic and phonological connectivity effects have been
extensively documented, little is known about their underlying
developmental mechanism. One possibility is that learning is driven by
lexical network growth where highly connected words in the child's early
lexicon enable learning of similar words. Another possibility is that
learning is driven by highly connected words in the external learning
environment, instead of highly connected words in the early internal
lexicon. The present study tests both scenarios systematically in both
the phonological and semantic domains across 10 languages. We show that
phonological and semantic connectivity in the learning environment
drives growth in both production- and comprehension-based vocabularies,
even controlling for word frequency and length. This pattern of findings
suggests a word learning process where children harness their
statistical learning abilities to detect and learn highly connected
words in the learning environment.}
  \keywords{Word learning; semantic network; phonological network; network growth;
cross-linguistic analysis. \\

@@

What factors shape vocabulary learning over the course of early
childhood? To investigate this question, scientists have adopted
multiple research strategies, from conducting controlled laboratory
experiments (e.g. Markman, 1990) to analyzing dense corpora capturing
language learning in context (e.g., B. C. Roy, Frank, DeCamp, Miller, \&
Roy, 2015). One prominent strategy consists in documenting the timeline
of words' acquisition, and studying the properties that make words easy
or hard to learn (e.g., J. C. Goodman, Dale, \& Li, 2008; Huttenlocher,
Haight, Bryk, Seltzer, \& Lyons, 1991). For example, J. C. Goodman et
al. (2008) found that, within a lexical category (e.g., nouns), higher
parental frequency is associated with earlier learning. Researchers have
studied the role of several other factors such as word length and the
mean length of utterances in which the word occurs (e.g., Braginsky,
Yurovsky, Marchman, \& Frank, in revision; Swingley \& Humphrey, 2018).

Besides word-level properties, the structure of the lexicon (that is,
how words relate to one another) is also linked to the the Age of
Acquisition (AoA) of words. The lexical structure can be characterized
in terms of a network where each node represents a word in the
vocabulary, and each link between two nodes represents a relationship
between the corresponding pair of words (e.g., Collins \& Loftus, 1975;
Luce \& Pisoni, 1998). Previous studies have investigated early
vocabulary structure by constructing networks using a variety of
word-word relations including shared semantic features (McRae, Cree,
Seidenberg, \& McNorgan, 2005), target-cue relationships in free
association norms (Nelson, McEvoy, \& Schreiber, 1998), co-occurrence in
child directed speech (MacWhinney, 2014), and phonological relatedness
(Vitevitch, 2008). These studies have generally found that children tend
to produce words that have higher neighborhood density (i.e., high
connectivity in the network) earlier, both at the phonological and the
semantic level (Carlson, Sonderegger, \& Bane, 2014; Hills, Maouene,
Riordan, \& Smith, 2010; Hills, Maouene, Maouene, Sheya, \& Smith, 2009;
Stella, Beckage, \& Brede, 2017; Storkel, 2009).

@@

\caption{Illustration of the two growth scenarios. Filled grey circles represent known words (Internal) at a certain point in time. The empty, numbered circles represent words that have not yet been learned (External) and which are candidates to enter the lexicon next. The identity of the word that is going to be learned depends on the growth scenario.  Here the squares indicate the node that drives growth in each scenario and the hand pointer indicates which word is likely to be learned. For INT, the utility of a candidate, external node is the average degree (i.e., number of links) of the internal nodes that it would attach to. Thus, according to INT, node 1 is more likely to enter the lexicon. For EXT, the utility of a candidate node is its degree in the entire network. According to EXT, node 2 is more likely to enter the lexicon next. This figure is based on an example from Hills et al (2009).}\label{fig:growth}
\end{figure}

Hills et al. (2009) investigated the growth of lexico-semantic networks
in toddlers and found that growth did not proceed according to INT as
was originally hypothesized by Steyvers and Tenenbaum (2005), but rather
according to
EXT.\footnote{Besides INT and EXT, the authors tested a third mechanism (called the lure of associates) which resembles EXT in that it is driven by the connectivity of external nodes, except that this connectivity is computed with respect to words that are known. However, EXT is the externally-driven scenario that best predicted the data in this previous work.}
This is an important finding because it suggests that learning in the
early stages is mostly driven by properties of the external input,
regardless of how past knowledge is organized. However, this work
explored the INT/EXT growth in a special case: networks that were based
on 1) semantic associations, 2) production-based vocabularies, and 3)
data from English-learning children, only. The extent to which this
result depends on the domain (e.g., semantic vs.~phonological
connectivity), the vocabulary measure (production vs.~comprehension) and
culture/language is thus an open area for investigation (Hills \& Siew,
2018). In this work, we test the generality of prior findings along
these three dimensions.

First, we study the phonological network in addition to the semantic
network. These two networks represent different ways the mental lexicon
is structured (Beckage \& Colunga, 2016). In particular, words that are
neighbors in the semantic network (e.g., \enquote{cat}, \enquote{dog})
are not necessarily neighbors in the phonological network, and vice
versa. Does the phonological network also predict word learning?
Previous work has found an effect of words' connectivity in the
phonological network on their age of learning (Carlson et al., 2014;
Stella et al., 2017; Storkel, 2009). In other words, words learned
earlier in life tend to sound similar to many other words than a word
learned later in life. However, this finding is \emph{a priori}
compatible with both INT and EXT, and previous studies did not
explicitly compare these two mechanisms. Here, we investigate whether
phonological networks, like semantic networks, grow through EXT, or if
they rather grow via INT (Figure~\ref{fig:growth}).

Second, we study vocabularies measured using both comprehension and
production. Previous studies have found differences between these
vocabularies in terms of their content and rate of acquisition (Bates,
Dale, \& Thal, 1995; Benedict, 1979; Fenson et al., 1994). These
differences may reflect the fact that comprehension and production do
not share the same constraints. For instance, whereas comprehension
depends on the ease with which words are stored and accessed, production
depends, additionally, on the ease with which words are articulated,
e.g., shorter words are produced earlier (Braginsky et al., in
revision). By investigating comprehension-based vocabularies, we assess
the extent to which the network growth mechanism captures general
learning patterns beyond the specific constraints of production.

Finally, we use developmental data in 10 languages. Lexical networks can
show more or less cross-linguistic variability along both the semantic
and phonological domains (Arbesman, Strogatz, \& Vitevitch, 2010; Lupyan
\& Lewis, 2017; Youn et al., 2016). Besides, cultures might differ in
the way caregivers talk to children (Cristia, Dupoux, Gurven, \&
Stieglitz, 2017; Kuhl et al., 1997), and this difference in the input
could influence the way in which the children's networks grow. Thus,
cross-linguistic comparison is crucial to test the extent to which
growth mechanisms are equally engaged across a wider variety of cultures
compared with the extent to which the growth mechanisms are specific to
patterns of learning that emerge due to the particulars of a given
language or culture (Bates \& MacWhinney, 1987; Slobin, 2014).

We adopted the following research strategy. We used parent reports on
the MacArthur-Bates Communicative Development Inventory and its
cross-linguistic adaptations (Fenson et al., 1994; Frank, Braginsky,
Yurovsky, \& Marchman, 2017). We studied the timeline of word learning
using the normative age of acquisition (i.e., the age at which at least
50\% of children know a given words). Our choice of studying the
normative learning trajectory rather than the individual trajectories
was motivated by the nature of the dataset used---which is primarily
based on cross-sectional studies. Children may vary in their individual
learning trajectories, but the aggregate data provide highly robust
measures of the average learning patterns (Fenson et al., 1994). The use
of such measures has lead to important insights on the mechanisms of
word learning (J. C. Goodman et al., 2008; Hills et al., 2010, 2009;
Stella et al., 2017; Storkel, 2009).

The paper is organized as follows. First, we describe the datasets we
used and explain how we constructed the networks. Second, we analyze
static properties of words' connectivity in these networks (correlation
with age of acquisition and shape of the distribution) and we explain
how these properties inform hypotheses about network growth. Next, we
fit the two hypothesized growth mechanisms to the data. We investigate
the extent to which the results obtained in Hills et al. (2009)
generalize to phonological networks and comprehension-based
vocabularies, and whether this generalization holds
cross-linguistically.

@@

For each word in the CDI data, we compute the proportion of children who
understand or produce the word at each month. Then we fit a logistic
curve to these proportions and determined when the curve crosses 0.5,
i.e., the age at which at least 50\% of children know the word. We take
this point in time to be each word's age of acquisition (Braginsky et
al., in revision; J. C. Goodman et al., 2008).

@@

\caption{\label{tab:stats}Statistics for the dataset we used.}
\centering
\begin{tabular}[t]{>{\bfseries}lrlrrlr}
\toprule
\multicolumn{1}{c}{} & \multicolumn{3}{c}{Comprehension} & \multicolumn{3}{c}{Production} \\
\cmidrule(l{2pt}r{2pt}){2-4} \cmidrule(l{2pt}r{2pt}){5-7}
Language & Nouns & Ages & N & Nouns & Ages & N\\
\midrule
Croatian & 209 & 8-16 & 250 & 312 & 16-30 & 377\\
Danish & 200 & 8-20 & 2,398 & 316 & 16-36 & 3,714\\
English & 209 & 8-18 & 2,435 & 312 & 16-30 & 5,520\\
French & 197 & 8-16 & 537 & 307 & 16-30 & 827\\
Italian & 209 & 7-24 & 648 & 312 & 18-36 & 752\\
Norwegian & 193 & 8-20 & 2,922 & 316 & 16-36 & 9,303\\
Russian & 207 & 8-18 & 768 & 314 & 18-36 & 1,037\\
Spanish & 208 & 8-18 & 788 & 312 & 16-30 & 1,146\\
Swedish & 205 & 8-16 & 467 & 339 & 16-28 & 900\\
Turkish & 180 & 8-16 & 1,115 & 297 & 16-36 & 2,422\\
\bottomrule
\end{tabular}
\end{table}

@@

Since the free association norms are available only in English, we used
the hand-checked translation equivalents available in Wordbank, which
allowed us to use the English association norms across 10 languages.
Semantic associations are not necessarily shared across languages, but
we use this technique as a reasonable first approximation. In support of
this approximation, Youn et al. (2016) showed that semantic networks
across languages share substantial similarities. Furthermore, using the
same association data across languages does not necessarily mean that
the resulting networks will grow in a similar fashion. For instance, the
set of words acquired by children as well as the order of word
acquisition can vary from language to language leading to possibility
different learning strategies.

@@

In previous studies, two nodes were linked if they had an edit distance
of 1 (Carlson et al., 2014; Stella et al., 2017; Storkel, 2009).
However, these studies reported a contribution of phonological
connectivity to noun learning when networks were built using a rich
adult vocabulary. Since the focus of the current study is on the
mechanism of growth, the networks should be based on children's early
vocabulary. The latter, however, contains very few noun pairs with an
edit distance of 1. Thus, we increased the threshold from 1 to 2, that
is, two nodes were related if their edit distance was equal to 1 or
2.\footnote{We also considered the case of an edit distance of 1 as well as the continuous measure, i.e., the inverse edit distance without threshold. In both cases, the results were weaker than those obtained with a threshold of 2. We did not consider the case of a threshold larger than 2 since many short pairs appear phonologically unrelated when the edit distance is 3 or more (e.g., "cat"/"dog").}
The connectivity of a given node was characterized with its
\emph{degree}: the number of links it shares with other words.

@@

We start by analyzing word connectivity in the global (static) network.
We constructed this network using nouns learned by the oldest age for
which we have CDI data (e.g., in English this corresponds, in
comprehension, to the network by 18 months, and in production, to the
network by 30 months). This global network is the end-state towards
which both INT and EXT converge by the last month of learning. Moreover,
following Hills et al. (2009), we used this end-state network as a proxy
for the external connectivity in the learning environment. Below we
analyze properties of these global networks that may a priori hint at an
INT- or EXT-like growth.

\begin{figure}[!h]
\includegraphics[width=\textwidth]{ms_files/figure-latex/corrProd-1} \caption{Production data (Age of acquisition) as predicted by the degree (i.e., connectivity) in this network. Results are shown in each language for phonological and semantic networks. Each point is a word, with lines indicating linear model fits, and numbers indicating the Pearson correlation coefficients.}\label{fig:corrProd}
\end{figure}

\begin{figure}[!h]
\includegraphics[width=\textwidth]{ms_files/figure-latex/corrComp-1} \caption{Comprehension data (Age of acquisition) as predicted by the degree (i.e., connectivity) in this network. Results are shown in each language for phonological and semantic networks. Each point is a word, with lines indicating linear model fits, and numbers indicating the Pearson correlation coefficients.}\label{fig:corrComp}
\end{figure}

@@

Figures \ref{fig:corrProd} and \ref{fig:corrComp} show how the age of
acquisition in production and comprehension, respectively, correlates
with the degree (or indegree for the semantic networks). For ease of
visual comparison, the predictor (i.e., the degree) was centered and
scaled. The plots show, overall, a negative correlation between the
month of acquisition and the degree. In production data, the average
correlation across languages was -0.24 (\(SD=\) 0.10) for the semantic
networks and -0.30 (\(SD=\) 0.05) for the phonological networks. In
comprehension data, the average correlation was -0.21 (\(SD=\) 0.08) for
the semantic networks and -0.21 (\(SD=\) 0.07) for the phonological
networks. These results indicate that nouns with higher degrees are
generally learned earlier, thus replicating previous findings in English
(Hills et al., 2009; Storkel, 2009) and extending these findings to 10
different languages, generally, in both production- and
comprehension-based vocabularies.

@@

We estimated the parameter \(\beta\) using a Bayesian approach. The
inference was performed using the probabilistic programming language
WebPPL (N. Goodman \& Stuhlmuller, 2014). We defined a uniform prior
over \(\beta\), and at each month, we computed the likelihood function
over words that could possibly enter the lexicon at that month, fit to
the words that have been learned at that month (using Formula 1). Markov
Chain Monte Carlo sampling resulted in a posterior distribution over
\(\beta\), which we summarized in Figure \ref{fig:growthPred}. The
results replicate Hills et al.'s original finding regarding the semantic
network in English and the production-based vocabulary, which is that
this network grows by EXT, not by INT. Crucially, our results show that,
generally speaking, this finding generalizes to comprehension-based
vocabulary, and holds across languages. This generalization was obtained
in both the semantic\footnote{One could imagine that the fact of using
  English free association norms cross-linguistically would decrease the
  effect of non-English semantic networks because of possible cultural
  differences. However, our findings do not support this assumption,
  rather, it supports our initial approximation about the shared nature
  of the semantic similarity measure. That said, this approximation is
  not perfect. For example there is evidence that a small part of the
  variance in free association data can be explained by phonological
  similarity (Kachergis, Cox, \& Jones, 2011; Matusevych \& Stevenson,
  2018), thus leading to possibly minor cross-linguistic differences.}
and phonological domains.

@@

Above we showed that the way semantic and phonological information is
structured in the learning environment contributes to noun learning (via
EXT) across languages. However, we know that other factors influence
learning as well (e.g., Braginsky et al., in revision). Next we
investigated how semantic and phonological connectivity interact with
two other factors. The first one is word frequency, a well studied
factor shown to predict the age of acquisition in a reliable fashion
(e.g. J. C. Goodman et al., 2008). The second factor is word length,
which was shown to correlate with phonological connectivity: Shorter
words are more likely to have higher connectivity (Pisoni, Nusbaum,
Luce, \& Slowiaczek, 1985; Vitevitch \& Rodríguez, 2005).

@@

For word length, we counted the number of phonemes in our generated IPA
transcription. For word frequency, we used the frequency estimates from
Braginsky et al. (in revision) where unigram counts were derived based
on CHILDES corpora in each language (MacWhinney, 2014). Although these
frequency counts use transcripts from independent sets of children, they
are based on large samples, and this allows us to average out possible
differences between children and the specificities of their input (see
J. C. Goodman et al., 2008 for a similar research strategy).

@@

The findings for the new predictors were as follows. Overall, frequency
is the largest and most consistent predictor of age of acquisition in
both comprehension and production data and across languages, endorsing
results for nouns across a variety of analyses (Braginsky et al., in
revision; J. C. Goodman et al., 2008; B. C. Roy et al., 2015). Word
length is more predictive for production than comprehension (and this
difference is very clear in the global model), replicating previous work
(Braginsky et al., in revision). Thus, word length seems to reflect the
effects of production's constraints rather than comprehension's
constraints, i.e., longer words are harder to articulate but they may
not be significantly more difficult to store and access.

@@

We found that the original findings reported in Hills et al. (2009)
generalize well across all these dimensions. First, just like semantic
networks, phonological networks grow via the externally-driven scenario
(EXT), not by the internally-driven mechanism (INT). Second,
comprehension-based vocabularies grow in a way similar to
production-based vocabularies. Finally, the findings were, overall,
similar across the 10 languages we tested. Although we find some
cross-linguistic variation when semantic and phonological networks were
pitted against frequency and length, this variability is to be taken
with a grain of salt as it might be exaggerated in our study by several
factors such as the limited and partially-overlapping set of nouns for
each language, measurement error due to the sample of acquisition data,
the sample of frequency data, and the translation of association norms.
In fact, both phonological and semantic connectivity are significant
predictors above and beyond frequency and length when data are pooled
across languages.

@@

For example, in the semantic domain, growth according to EXT could be
explained by a mechanism similar to cross-situational learning (McMurray
et al., 2012; Smith \& Yu, 2008; Yurovsky \& Frank, 2015). According to
this mechanism, children track the co-occurrence of concrete nouns with
their possible semantic referents. The referent of a word heard in only
one naming situation can be ambiguous (e.g., when the word
\enquote{ball} is heard for the first time in the presence of both a
ball and a chair), but hearing the same word in a diversity of semantic
contexts allows the learner to narrow down the set of possible
word-object mappings. In our case, free association (used to determine
semantic network connectivity) is related to contextual co-occurrence
(Fourtassi \& Dupoux, 2013; Griffiths, Steyvers, \& Tenenbaum, 2007),
meaning that highly connected words will tend to occur in a variety of
speech and referential contexts. This fact makes such words easier to
learn because they have more referential disambiguating cues across
learning contexts. Crucially, children can learn these words without
necessarily knowing the meaning of all other words with which they
co-occur (hence the similarity with EXT). This possibility is supported
by the finding that words' diversity of occurrence in child directed
speech predicts their age of learning (Hills et al., 2010; Stella et
al., 2017).

@@

\hypertarget{ref-beckage}{}
Beckage, N. M., \& Colunga, E. (under review). Network growth modeling
to capture individual lexical learning.

@@

\hypertarget{ref-hills2018}{}
Hills, T. T., \& Siew, C. S. (2018). Filling gaps in early word
learning. \emph{Nature Human Behaviour}, \emph{2}(9).
